---
title: Text Mining Clustering Tweets
author: andrea perlato
date: '2019-08-02'
slug: text-mining-clustering-tweets
categories:
  - text mining
tags:
  - clustering
  - hierarchical clustering
---

<style>
body {
text-align: justify}
</style>

We can use Unsupervised Classification (clustering) to learn more about text that we have to analyze. More specifically, we use [**Hierarchical clustering**](https://www.andreaperlato.com/theorypost/introduction-to-k-means-and-hierarchical-clustering/) to cluster our text into groups that are the propensity to occur together. In this example, text are some tweet about [**Catalan Independence Referendum**](https://en.wikipedia.org/wiki/2017_Catalan_independence_referendum).

```{r, echo=FALSE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
setwd("C:/07 - R Website/dataset/ML")
d=read.csv("Catalan.csv")

```

First step is to convert our tweets into a corpus, remove punctuation, transform everything into lowercase, remove numbers, and stop words. Then, we have to stamming the document, and finally we have a corpus of terms. with out sparse terms. 

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
library(tm)

corpus <- Corpus(VectorSource(d$text))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
corpus <- tm_map(corpus, stemDocument)

# Create term document matrix
tdm <- TermDocumentMatrix(corpus,control = list(minWordLength=c(1,Inf)))

t <- removeSparseTerms(tdm, sparse=0.90) # remove words not appear 90% of the time

```

Next step is to create a matrix of word frequency, and we can plot the words that occur more than 200 times, as shown in the graph below.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
m <- as.matrix(t) # create a matrix of word frequencies

# Plot frequent terms
freq <- rowSums(m)
freq <- subset(freq, freq>=200)

barplot(freq, las=2, col = rainbow(25))

```

Obviously, **catalonia** is the most common term, but we have also other interesting high frequency words, such as **solidar**.
Now, we can use Hierarchical clustering of the words using Dendrogram.

```{r, echo=TRUE, comment=NA, message=FALSE, warning=FALSE, fig.align='center', out.width = '100%'}
# Hierarchical word tweet clustering using dendrogram 

distance <- dist(scale(m)) # distance between different terms

# print(distance, digits = 2)

hc <- hclust(distance, method = "ward.D") # use ward.D for heirarchial clustering
# plot(hc, hang=-1)
# rect.hclust(hc, k=12)

library(dendextend) # color dendrogram

dend <- hc
dend <- color_branches(dend, k = 3)
dend <- color_labels(dend, k = 3)

# Represent the different  clusters with different colors
plot(dend, main = 'Cluster Dendrogram', ylab = 'Height')

```

From the graph above, we can see that to different cluster has been assigned a different color.
To learn more about Hierarchical clustering and Dendrogram, [**look at this post**](https://www.andreaperlato.com/theorypost/introduction-to-k-means-and-hierarchical-clustering/).

















































